---
title: "PGP_assays"
author: "Shawn Higdon"
date: "11/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# libraries
```{r}
library(dplyr)
```


## Read in the Data for PGP Biochemical assays

```{r, message=FALSE}
# Create a list of BCW isolates
bcw_isolate_list <- read.table("bcw_isolate_list.txt")

# Read in the hmmscan Search output files for nif genes in each isolate's annotated faa list
pgp_data_list <- list.files(path = "./PGP_Data", pattern = 'PGP_*', recursive = T, full.names = T)
pgp_data_list

# Create a list of bcw nifscan search output tables
pgp_data_tbl_list <- lapply(pgp_data_list, read_csv)

# create object for each PGP Assay
pgp_accd <- pgp_data_tbl_list[[1]]
pgp_auxin <- pgp_data_tbl_list[[2]]
pgp_po4 <- pgp_data_tbl_list[[3]]
pgp_sidero <- pgp_data_tbl_list[[4]]

```


## Compute Assay metrics

### ACC Deaminase Growth Assay

> Isolates cultured on medium containing 1-amino-1-cyclopropane carboxylic acid as the only source of available nitrogen

```{r}
# ACCD Assay
str(pgp_accd)
## Compute average for Day 0 OD600 replicates
pgp_accd$OD600_D0_avg <- rowMeans(pgp_accd[,3:5], na.rm = TRUE)
## Compute average for Day 4 OD600 replicates
pgp_accd$OD600_D4_avg <- rowMeans(pgp_accd[,6:8], na.rm = TRUE)

## Compute Relative growth rate for each isolate
pgp_accd$ACCD_RGR <- pgp_accd$OD600_D4_avg/pgp_accd$OD600_D0_avg

## change number of significant digits for numeric values to 3
pgp_accd %>% mutate_if(is.numeric, round, digits = 3)


```

### Auxin Biosynthesis Assay

```{r}
str(pgp_auxin)

## Compute average for OD535 replicates
pgp_auxin$OD535_avg <- rowMeans(pgp_auxin[,3:5], na.rm = TRUE)

## Compute average for auxin concentration replicate values
pgp_auxin$IAA_mg_ml_avg <- rowMeans(pgp_auxin[,6:8], na.rm = TRUE)

## change number of significant digits for numeric values to 3
pgp_auxin %>% mutate_if(is.numeric, round, digits = 3)
```

## Phosphate Solubilization Assay

```{r}
str(pgp_po4)

## Compute average for OD535 replicates
pgp_po4$OD535_avg <- rowMeans(pgp_po4[,3:5], na.rm = TRUE)

## Compute average for auxin concentration replicate values
pgp_po4$PO4_mg_L_avg <- rowMeans(pgp_po4[,6:8], na.rm = TRUE)

## Multiply diluted concentrations by assay dilution factor of 100
pgp_po4$PO4_mg_L_avg <- pgp_po4$PO4_mg_L_avg*100

## change number of significant digits for numeric values to 3
pgp_po4 %>% mutate_if(is.numeric, round, digits = 3)
```

## Siderophore Production Assay

```{r}
str(pgp_sidero)

## Compute average for OD535 replicates
pgp_sidero$OD630_avg <- rowMeans(pgp_sidero[,3:5], na.rm = TRUE)

## Compute average for Siderophore units (%) using AR_AS replicate values
## % Siderophore units = Absorbance of Sample (AS) - Absorbance of Reference (AR) / AR
pgp_sidero$Sidero_units_avg <- rowMeans(pgp_sidero[,6:8], na.rm = TRUE)

## change number of significant digits for numeric values to 3
pgp_sidero %>% mutate_if(is.numeric, round, digits = 3)
```

## Read in 15N Assay data

> Bring in 15N/14N ratios from 15N assimilation assay

```{r}
# Read in csv of 15N/14N ratio for each isolate
N15_ratios <- read.csv("./../15N_assay_data_SH.csv", header = T)

```

## Combine PGP Assay Results

> Make a data frame with results from:

1. 15N Assay

2. ACCD Growth Assay

3. Auxin Biosynthesis Assay

4. Phosphate Solubilization Assay

5. Siderophore Production Assay

### Make the Data Frame
```{r}
# Read in asm_no_euk_stats
asm_no_euk_stats <- read.csv("./asm_stats_no_euk.csv", header = T)

# start with 15N Assay ratios
pgp_master <- tibble(
  BCW_ID=asm_no_euk_stats$Isolate)

# add N15 ratio values for each isolate
pgp_master$`BNF` <- N15_ratios$ratio[match(pgp_master$BCW_ID, N15_ratios$BCW_ID)]

# add ACCD assay results
pgp_master$`ACC` <- pgp_accd$ACCD_RGR[match(pgp_master$BCW_ID, pgp_accd$BCW_ID)]

# add Auxin assay results
pgp_master$`IAA` <- pgp_auxin$IAA_mg_ml_avg[match(pgp_master$BCW_ID, pgp_auxin$BCW_ID)]

# add Phosphate solubility assay results
pgp_master$`PO4` <- pgp_po4$PO4_mg_L_avg[match(pgp_master$BCW_ID, pgp_po4$BCW_ID)]

# add siderophore assay results
pgp_master$`SID` <- pgp_sidero$Sidero_units_avg[match(pgp_master$BCW_ID, pgp_sidero$BCW_ID)]

# read in bcw_id and genus map table
bcw_id_genus <- read.csv("./bcw_id_genus.csv", header = T)

# Genus classification for each isolate
pgp_master$Genus <- bcw_id_genus$Genus[match(pgp_master$BCW_ID, bcw_id_genus$BCW_ID)]

## assign a unique color to each genus
pgp_genus_colors <- data.frame(Genus = unique(pgp_master$Genus), Color =  distinctColorPalette(length(unique(pgp_master$Genus))))
### match
pgp_master$Color <- pgp_genus_colors$Color[match(pgp_master$Genus, pgp_genus_colors$Genus)]

# add genome assembly size
pgp_master$Assembly_size <- asm_no_euk_stats$Total.Length..Mb.[match(pgp_master$BCW_ID, asm_no_euk_stats$Isolate)]

# add n_bin data
pgp_master$n_bins <- asm_no_euk_stats$Bins..n.[match(pgp_master$BCW_ID, asm_no_euk_stats$Isolate)]

## change number of significant digits for numeric values to 3
pgp_master <- pgp_master %>% mutate_if(is.numeric, round, digits = 3)

# Remove incomplete cases (Isolates with NA values for PGP assays)
pgp_master <- drop_na(pgp_master)
str(pgp_master)

```

### Complex Heatmap for All PGP isolate assay data
```{r}
# Create matrices for plotting & make rownames isolate id

## 15N Ratio
pgp_master_N15_mat <- as.matrix(pgp_master[,2])
rownames(pgp_master_N15_mat) <- pgp_master$BCW_ID
pgp_N15_col <- inferno(direction = -1, 6)

## ACCD Assay
pgp_master_ACCD_mat <- as.matrix(pgp_master[,3])
rownames(pgp_master_ACCD_mat) <- pgp_master$BCW_ID
pgp_ACCD_col <- viridis(direction = -1, 6)

## Auxin Assay
pgp_master_auxin_mat <- as.matrix(pgp_master[,4])
rownames(pgp_master_auxin_mat) <- pgp_master$BCW_ID
pgp_auxin_col <- plasma(direction = -1, 6)

## Phosphate Assay
pgp_master_po4_mat <- as.matrix(pgp_master[,5])
rownames(pgp_master_po4_mat) <- pgp_master$BCW_ID
pgp_po4_col <- magma(direction = -1, 6)

## Siderophore Assay
pgp_master_Siderophore_mat <- as.matrix(pgp_master[,6])
rownames(pgp_master_Siderophore_mat) <- pgp_master$BCW_ID
pgp_Siderophore_col <- cividis(direction = -1, 6)



## 15N Heatmap
pgp_master_N15_chm <- 
  Heatmap(pgp_master_N15_mat,
          name = "BNF (15N/14N)",
          show_row_names = F,
          col = pgp_N15_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 5, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## ACCD RGR Heatmap
pgp_master_ACCD_chm <- 
  Heatmap(pgp_master_ACCD_mat,
          name = "ACC (RGR)",
          show_row_names = F,
          col = pgp_ACCD_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 5, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Auxin Heatmap
pgp_master_auxin_chm <- 
  Heatmap(pgp_master_auxin_mat,
          name = "IAA (mg/mL)",
          show_row_names = F,
          col = pgp_auxin_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 220, 50),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Phosphate Heatmap
pgp_master_po4_chm <- 
  Heatmap(pgp_master_po4_mat,
          name = "PO4 (mg/L)",
          show_row_names = F,
          col = pgp_po4_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 2200, 500),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Siderophore Heatmap
pgp_master_Siderophore_chm <- 
  Heatmap(pgp_master_Siderophore_mat,
          name = "SID (% Units)",
          show_row_names = F,
          col = pgp_Siderophore_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(-1.5, 1.0, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Genus Heatmap
pgp_genus_col <- unique(pgp_master$Color)

pgp_master_genus_chm <-
  Heatmap(pgp_master$Genus,
          name = "Genus",
          col = pgp_genus_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(ncol = 2,
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          width = unit(5, "mm"))

pgp_master_list <- pgp_master_N15_chm + 
  pgp_master_ACCD_chm + 
  pgp_master_auxin_chm + 
  pgp_master_po4_chm +
  pgp_master_Siderophore_chm +
  pgp_master_genus_chm


pgp_master_chm <- 
draw(pgp_master_list, heatmap_legend_side = "right")

```


### Select Single Isolates

```{r}
# Filter to select only genus classified isolates (pure)
pgp_master_single_ids <- pgp_master %>% filter(Genus !="unassigned")
#                                               &
#                                      n_bins != "2" &
#                                      n_bins != "3" &
#                                      n_bins != "4" &
#                                      n_bins != "5")
#
# Remove incomplete cases (Isolates with NA values for PGP assays)
pgp_master_single_ids <- drop_na(pgp_master_single_ids)

# Genus colors for single isolate cases
pgp_genus_single_col <- unique(pgp_master_single_ids$Color)

# Create matrices for plotting & make rownames isolate id

## 15N Ratio
pgp_master_single_ids_N15_mat <- as.matrix(pgp_master_single_ids[,2])
rownames(pgp_master_single_ids_N15_mat) <- pgp_master_single_ids$BCW_ID
pgp_N15_col <- inferno(direction = -1, 6)

## ACCD Assay
pgp_master_single_ids_ACCD_mat <- as.matrix(pgp_master_single_ids[,3])
rownames(pgp_master_single_ids_ACCD_mat) <- pgp_master_single_ids$BCW_ID
pgp_ACCD_col <- viridis(direction = -1, 6)

## Auxin Assay
pgp_master_single_ids_auxin_mat <- as.matrix(pgp_master_single_ids[,4])
rownames(pgp_master_single_ids_auxin_mat) <- pgp_master_single_ids$BCW_ID
pgp_auxin_col <- plasma(direction = -1, 6)

## Phosphate Assay
pgp_master_single_ids_po4_mat <- as.matrix(pgp_master_single_ids[,5])
rownames(pgp_master_single_ids_po4_mat) <- pgp_master_single_ids$BCW_ID
pgp_po4_col <- magma(direction = -1, 6)

## Siderophore Assay
pgp_master_single_ids_Siderophore_mat <- as.matrix(pgp_master_single_ids[,6])
rownames(pgp_master_single_ids_Siderophore_mat) <- pgp_master_single_ids$BCW_ID
pgp_Siderophore_col <- cividis(direction = -1, 6)



## 15N Heatmap
pgp_master_single_ids_N15_chm <- 
  Heatmap(pgp_master_single_ids_N15_mat,
          name = "BNF (15N/14N)",
          width = unit(0, "cm"),
          show_row_names = F,
          col = pgp_N15_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 5, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## ACCD RGR Heatmap
pgp_master_single_ids_ACCD_chm <- 
  Heatmap(pgp_master_single_ids_ACCD_mat,
          name = "ACC (RGR)",
          show_row_names = F,
          col = pgp_ACCD_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 5, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Auxin Heatmap
pgp_master_single_ids_auxin_chm <- 
  Heatmap(pgp_master_single_ids_auxin_mat,
          name = "IAA (mg/mL)",
          show_row_names = F,
          col = pgp_auxin_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 220, 50),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Phosphate Heatmap
pgp_master_single_ids_po4_chm <- 
  Heatmap(pgp_master_single_ids_po4_mat,
          name = "PO4 (mg/L)",
          show_row_names = F,
          col = pgp_po4_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(0, 2200, 500),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Siderophore Heatmap
pgp_master_single_ids_Siderophore_chm <- 
  Heatmap(pgp_master_single_ids_Siderophore_mat,
          name = "SID (% Units)",
          show_row_names = F,
          col = pgp_Siderophore_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(at = seq(-1.5, 1.0, 0.5),
                                      legend_direction = "horizontal",
                                      legend_width = unit(6, "cm"),
                                      color_bar = "continuous",
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          row_names_gp = gpar(fontsize = 1.2))

## Genus Heatmap
pgp_genus_col <- unique(pgp_master_single_ids$Color)

pgp_master_single_ids_genus_chm <-
  Heatmap(pgp_master_single_ids$Genus,
          name = "Genus",
          col = pgp_genus_single_col,
          column_names_gp = gpar(fontsize = 10),
          heatmap_legend_param = list(ncol = 2,
                                      labels_gp = gpar(fontsize = 10),
                                      title_position = "topcenter"),
          width = unit(5, "mm"))

pgp_master_single_ids_list <- pgp_master_single_ids_N15_chm + 
  pgp_master_single_ids_ACCD_chm + 
  pgp_master_single_ids_auxin_chm + 
  pgp_master_single_ids_po4_chm +
  pgp_master_single_ids_Siderophore_chm +
  pgp_master_single_ids_genus_chm


pgp_master_single_ids_chm <- 
draw(pgp_master_single_ids_list, ht_gap = unit(0.1, "cm"), heatmap_legend_side = "right")


```












